% Thesis Acknowledgements ------------------------------------------------


%\begin{acknowledgementslong} %uncommenting this line, gives a different acknowledgements heading
\begin{acknowledgements}      %this creates the heading for the acknowlegments

Although I didn't know it yet, this journey began almost 10 years ago in Boston, with Susana Muñoz. A study about the Ivory Coast, a presentation at M.I.T. and a strong determination to continue learning. New plans and a new future. Oscar Corcho then appeared, and thanks to him everything started to take shape. I discovered science in the Ontology Engineering Group, with great researchers who are also great colleagues. Asunción Gómez, Jorge Gracia, Victor Fernandez, Raul Garcia, Elena Montiel, Mariano Fernandez, Jose Angel Ramos and Mariano Rico. Particularly impressive were the talks with María Poveda, Dani Vila and Julia Bosque. New ideas and unforeseen approaches came up in every talk. I admire you. Research means reading, understanding, testing and also collaborating. I am very grateful for the interactions with Daniel Garijo, Filip Radulovic and Idafen Santana, with Pablo Calleja and Rafael Gonzalez, with Nandana Mihindukulasooriya and Freddy Priyatna, with Esteban Gonzalez and with Fernando Serena, and the dedicated support of Raul Alcazar and Miguel Angel, because you forced me to be more pragmatic. And I cannot forget the DiversiTeam, where all the extra skills are put into practice. I don't want to forget anyone, you are all part of this work: David Chaves, Paola Espinoza, Francisco Yedro, Alba Fernandez, Maria Navas, Patricia Martin, Elvira Amador, Edna Ruckhaus, Alvaro Brandon, Serge Chavez, Andres Prados, Ana Iglesias, and Julian Arenas.

Special mention for my supervisor, Oscar Corcho, and my mentor, Jose Luis Redondo. Both of you have the ability to choose the right words and simplify any task. You have complemented each other perfectly, even from the distance, and I have enjoyed every year of this thesis. Oscar knows how to touch the right keys at the right time, and Jose Luis reduces each problem to the challenge that motivates it, no matter how complex it was. I am extremely lucky to have you by my side on this journey, and and your influence will guide my future steps.

Thank you so much Beatriz for offering me so much help this time. These have been very intense years, with many changes and challenges, but we have been able to keep smiling and move forward. I hope Martin and Alonso feel proud of us for having achieved our objectives. I will miss our nights working with their crying in the background.

Federico and Loli, I had imagined it otherwise, but I know you will be proud of me. Not for the thesis, but for the decisions that led me to do it. You are in everything I do, and thanks to your effort this achievement has become a reality. I also have to thank the patience of my sisters, Ana and Lucia, who have always been able to push me forward.



\end{acknowledgements}
%\end{acknowledgementslong}

% ------------------------------------------------------------------------


